---
title: "Introduction to computePI"
author: "Samer Mouksassi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to computePI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
require(vpc)
require(rlang)
require(dplyr)
require(ggplot2)
require(data.table)
require(computePI)
```


When producing VPC plots you must:

* look at your observed data, understand the design of it and how best to plot it.

* Decide whether you want to stratify (e.g. by Study, by Dose).

* Decide whether you want to bin the x variable and how.

* Repeat the above on the simulated data.

* Compute the prediction intervals and their confidence intervals.

* Produce a plot that compare observed to simulated PI.

The compute PI package makes the computations of needed percentiles fast and easy:


* It uses efficient computations using data.table, so you spend less time waiting.


To get started:
```
devtools::install_github("smouksassi/computePI")
library(computePI)
```



## Observed data
Let us have a look at the observed data first overall and then zooming into Time > 120.
```{r ,fig.show='hold'}
ggplot(obsdata) +
  facet_wrap( ~ DVTYPE,scales = "free",ncol=3) +
  geom_point(aes(TIME,DV))+
  theme_bw(base_size = 10)

ggplot(obsdata[obsdata$TIME > 120,]) +
  facet_wrap( ~ DVTYPE,scales = "free",ncol=3) +
  geom_point(aes(TIME,DV))+
  theme_bw(base_size = 10)

```

## ComputePI Options
Let us focus on the data with Time > 120 and test the functionality of different number of binning by DVTYPE.

Data preparation
```{r}
obsdata <- obsdata %>%
  arrange(DVTYPE, ID, TIME)
simdata <- simdata %>%
  arrange(REPL, DVTYPE, ID, TIME)

obsdata <- obsdata[obsdata$TIME > 120, ]
simdata <- simdata[simdata$REPL < 20, ]
simdata <- simdata[simdata$TIME > 120, ]
nsims<- nrow(simdata)/nrow(obsdata)
obsdata <-obsdata %>%
  mutate(nbins= 
    case_when(
 DVTYPE=="CMObs" ~5,
 DVTYPE=="CObs" ~ 10,
 DVTYPE=="EObs" ~ 6,
 TRUE ~ 10
))
simdata$nbins <- rep(obsdata$nbins,nsims)

```

```{r}
VPCPIDATA<- computePI(
  obsdata = obsdata, simdata = simdata, stratify = ~DVTYPE,REP = REPL,
  NBINS = nbins, LLOQ = NULL)
```

```{r, echo=FALSE, fig.width=7}
AAA <- ggplot(VPCPIDATA$PI) +
  facet_grid( ~ DVTYPE) +
  geom_ribbon(
    aes(
      XMED,
      ymin = (`SIM2.5%CI`),
      ymax = (`SIM97.5%CI`),
      fill = QNAME,
      col = QNAME,
      group = QNAME
    ),
    alpha = 0.1,
    col = NA
  ) +
  geom_line(aes(
    XMED,
    y = `SIM50%CI`,
    col = QNAME,
    group = QNAME
  )) +
  geom_line(aes(
    x = XMED,
    y = RAWOBS,
    group = QNAME,
    linetype = QNAME
  ), size = 1) +
  geom_rug(data=VPCPIDATA$BINS ,aes(x= XMIN))+
  geom_rug(data=VPCPIDATA$BINS ,aes(x= XMAX))+
  geom_vline(data=VPCPIDATA$BINS,aes(xintercept= XMED) ,color ="gray")+
  
  scale_colour_manual(
    name = "Simulated Percentiles\nMedian (lines) 95% CI (areas)",
    breaks = c("5%PI", "50%PI", "95%PI", "Percent BLQ"),
    values = c("red", "blue", "red", "black"),
    labels = c("5%", "50%", "95%", "Percent BLQ")
  ) +
  scale_fill_manual(
    name = "Simulated Percentiles\nMedian (lines) 95% CI (areas)",
    breaks = c("5%PI", "50%PI", "95%PI", "Percent BLQ"),
    values = c("red", "blue", "red", "black"),
    labels = c("5%", "50%", "95%", "Percent BLQ")
  ) +
  scale_linetype_manual(
    name = "Observed Percentiles\n(black lines)",
    breaks = c("5%PI", "50%PI", "95%PI"),
    values = c("dotted", "solid", "dashed"),
    labels = c("5%", "50%", "95%")
  ) +
  guides(
    fill = guide_legend(order = 2),
    colour = guide_legend(order = 2),
    linetype = guide_legend(order = 1)
  ) +
  theme_bw(base_size = 10)+
  theme(
    legend.position = "top",
    legend.key.width = grid::unit(0.5, "cm"),
    axis.text.x = element_text(angle = 30)
  )
AAA + ylab("simulated/observed")+
   xlab("Time")

```

```{r, echo=FALSE, fig.width=7}
AAA <- ggplot(VPCPIDATA$PI) +
  facet_grid( ~ DVTYPE) +
  geom_rect(
    aes(
      xmin=XLEFT,
      xmax=XRIGHT,
      ymin = (`SIM2.5%CI`),
      ymax = (`SIM97.5%CI`),
      fill = QNAME,
      col = QNAME,
      group = QNAME
    ),
    alpha = 0.1,
    col = NA
  ) +
  geom_line(aes(
    XMED,
    y = `SIM50%CI`,
    col = QNAME,
    group = QNAME
  )) +
  geom_line(aes(
    x = XMED,
    y = RAWOBS,
    group = QNAME,
    linetype = QNAME
  ), size = 1) +
  geom_rug(data=VPCPIDATA$BINS ,aes(x= XMIN))+
  geom_rug(data=VPCPIDATA$BINS ,aes(x= XMAX))+
  geom_vline(data=VPCPIDATA$BINS,aes(xintercept= XMED) ,color ="gray")+
  
  scale_colour_manual(
    name = "Simulated Percentiles\nMedian (lines) 95% CI (areas)",
    breaks = c("5%PI", "50%PI", "95%PI", "Percent BLQ"),
    values = c("red", "blue", "red", "black"),
    labels = c("5%", "50%", "95%", "Percent BLQ")
  ) +
  scale_fill_manual(
    name = "Simulated Percentiles\nMedian (lines) 95% CI (areas)",
    breaks = c("5%PI", "50%PI", "95%PI", "Percent BLQ"),
    values = c("red", "blue", "red", "black"),
    labels = c("5%", "50%", "95%", "Percent BLQ")
  ) +
  scale_linetype_manual(
    name = "Observed Percentiles\n(black lines)",
    breaks = c("5%PI", "50%PI", "95%PI"),
    values = c("dotted", "solid", "dashed"),
    labels = c("5%", "50%", "95%")
  ) +
  guides(
    fill = guide_legend(order = 2),
    colour = guide_legend(order = 2),
    linetype = guide_legend(order = 1)
  ) +
  theme_bw(base_size = 10)+
  theme(
    legend.position = "top",
    legend.key.width = grid::unit(0.5, "cm"),
    axis.text.x = element_text(angle = 30)
  )
AAA + ylab("simulated/observed")+
   xlab("Time")+
  scale_x_log10()

```

